---
title: 'Figure Display Test'
author: 'AI Assistant'
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 2
    number_sections: true
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center')
library(dplyr)
library(ggplot2)
library(knitr)
```

# Figure Display Test

This document tests the display of representative figures from our miRNA SNV analysis.

## Test Results

```{r test-results}
test_figures <- data.frame(
  Figure_Name = c(
    "01_enhanced_preprocessing_pipeline.png",
    "02_enhanced_data_quality_metrics.png",
    "01_boxplot_edad_grupo.png",
    "01_boxplot_oxidative_score.png",
    "01_pca_scatter_pc1_pc2.png",
    "02_pca_scatter_pc1_pc3.png",
    "01_family_contributions_heatmap.png",
    "02_position_contributions.png"
  ),
  Analysis_Type = c(
    "Enhanced Styling",
    "Enhanced Styling",
    "General Analysis",
    "General Analysis",
    "PCA Analysis",
    "PCA Analysis",
    "Pathway Analysis",
    "Pathway Analysis"
  ),
  File_Exists = file.exists(file.path('test_figures', c(
    "01_enhanced_preprocessing_pipeline.png",
    "02_enhanced_data_quality_metrics.png",
    "01_boxplot_edad_grupo.png",
    "01_boxplot_oxidative_score.png",
    "01_pca_scatter_pc1_pc2.png",
    "02_pca_scatter_pc1_pc3.png",
    "01_family_contributions_heatmap.png",
    "02_position_contributions.png"
  ))),
  stringsAsFactors = FALSE
)

knitr::kable(test_figures, caption = 'Test Figure Status')
```

## 1. Enhanced Styling - 01_enhanced_preprocessing_pipeline.png

**File Type:** PNG

**Description:** Data preprocessing pipeline visualization

```{r fig-
1
, fig.cap='
01_enhanced_preprocessing_pipeline.png
'}
fig_path <- file.path('test_figures', '
01_enhanced_preprocessing_pipeline.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
01_enhanced_preprocessing_pipeline.png
</div>')
}
```

## 2. Enhanced Styling - 02_enhanced_data_quality_metrics.png

**File Type:** PNG

**Description:** Analysis visualization

```{r fig-
2
, fig.cap='
02_enhanced_data_quality_metrics.png
'}
fig_path <- file.path('test_figures', '
02_enhanced_data_quality_metrics.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
02_enhanced_data_quality_metrics.png
</div>')
}
```

## 3. General Analysis - 01_boxplot_edad_grupo.png

**File Type:** PNG

**Description:** Boxplot comparison between groups

```{r fig-
3
, fig.cap='
01_boxplot_edad_grupo.png
'}
fig_path <- file.path('test_figures', '
01_boxplot_edad_grupo.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
01_boxplot_edad_grupo.png
</div>')
}
```

## 4. General Analysis - 01_boxplot_oxidative_score.png

**File Type:** PNG

**Description:** Oxidative load comparison between groups

```{r fig-
4
, fig.cap='
01_boxplot_oxidative_score.png
'}
fig_path <- file.path('test_figures', '
01_boxplot_oxidative_score.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
01_boxplot_oxidative_score.png
</div>')
}
```

## 5. PCA Analysis - 01_pca_scatter_pc1_pc2.png

**File Type:** PNG

**Description:** Principal Component Analysis visualization

```{r fig-
5
, fig.cap='
01_pca_scatter_pc1_pc2.png
'}
fig_path <- file.path('test_figures', '
01_pca_scatter_pc1_pc2.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
01_pca_scatter_pc1_pc2.png
</div>')
}
```

## 6. PCA Analysis - 02_pca_scatter_pc1_pc3.png

**File Type:** PNG

**Description:** Principal Component Analysis visualization

```{r fig-
6
, fig.cap='
02_pca_scatter_pc1_pc3.png
'}
fig_path <- file.path('test_figures', '
02_pca_scatter_pc1_pc3.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
02_pca_scatter_pc1_pc3.png
</div>')
}
```

## 7. Pathway Analysis - 01_family_contributions_heatmap.png

**File Type:** PNG

**Description:** Heatmap visualization of data patterns

```{r fig-
7
, fig.cap='
01_family_contributions_heatmap.png
'}
fig_path <- file.path('test_figures', '
01_family_contributions_heatmap.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
01_family_contributions_heatmap.png
</div>')
}
```

## 8. Pathway Analysis - 02_position_contributions.png

**File Type:** PNG

**Description:** Analysis visualization

```{r fig-
8
, fig.cap='
02_position_contributions.png
'}
fig_path <- file.path('test_figures', '
02_position_contributions.png
')
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  cat('<div class="alert alert-danger">Figure not found: 
02_position_contributions.png
</div>')
}
```

# Test Summary

This test validates that figures can be properly displayed in HTML format.

## Next Steps

1. Verify all figures display correctly
2. Create working presentation with verified figures
3. Validate all figures display correctly in final presentation

---

*Generated on: `r format(Sys.Date(), "%Y-%m-%d")`*
